Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-08 paper/sec-analysis.tex
more work on the text

>>>>

We now study the discrimination power to separate signal
and background events contained in the jet substructure
variables.
%
In Fig.~\ref{fig:mva_substructure_1}
we show the distributions of some of
these substructure variables for the boosted category: the
$k_t$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}), for
the subleading Higgs,
the energy correlation double ratio $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}), for the leading Higgs,
and then 
the 2--to--1 subjettiness ratio $\tau_{12}$,
Eq.~(\ref{eq:tau21}),
for both the leading and subleading Higgs candidates.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{plots/split12_h1_C1_boost.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h1_C1_boost.pdf}
  \caption{\small Distribution of representative substructure variables
    in the boosted category at the end of the cut-based
    analysis, used as input to the MVA.
    %
    From top to bottom and from left to right we show  the
$k_t$ splitting scale for
the subleading Higgs, the subjettiness ratio $\tau_{12}$
for the leading Higgs,
and the energy correlation double ratio $D_2$
for the leading and subleading Higgs.
%
}
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Fig.~\ref{fig:mva_substructure_1}
we can observe that for these substructure variables the shapes of the signal
and background distributions show important differences,
reflecting the inherent differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.
%
Importantly, we observe that signal and background distributions peak
in different regions: for example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (50 GeV) for signal (background) events, while
the ECF ratio $D_2^{(\beta)}$ peaks at much smaller
values for signal than
for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are similar
in the leading and in the subleading jets, therefore the information
contained in
$\tau_{12}$ from each of the two Higgs candidates  will be to some
extent redundant.

Therefore, while we don't explicitely have cuts in the substructure
variables, their discrimination power between signal and
background events is the motivation to include them
as input to the MVA, as we will explain in Sect.~\ref{sec:mva}.

### 2015-12-08 paper/sec-mva.tex
more work on the text

From Fig.~\ref{fig:exampleroc} we see that, always starting from the signal
and background events that survive the basic cut-based analysis,
one can achieve a  signal
efficiency of $\sim 90\%$ while reducing the background by
$80\%$, $60\%$ and $50\%$ in the boosted, intermediate and resolved
categories respectively.
%
It is easy to translate these factors into an  improvement factor of the pre-MVA
signal significance $S/\sqrt{B}$.
%
As an illustration, if a signal efficiency of 90\%
is required, we find that the MVA improves $S/\sqrt{B}$
by  approximately 2.0, 1.4 and 1.3  in the three categories respectively.

As mentioned above, an
 important restriction on the range of allowed cuts in the ANN output
 is that $y_{\rm cut}$
 cannot be so hard that the actual number of signal events expected
at the HL-LHC becomes unreasonably small.
%
To verify how many signal and background events are left after the MVA cut,
we show in Fig.~\ref{fig:nev2} the number of signal (dashed lines) and background (solid lines)
  events expected at the HL-LHC as a function of $y_{\rm cut}$,
  for the three exclusive categories.
  %
  Note that the value of the cut in the ANN cut is separately optimised in the three
  categories.

As can be seen from Fig.~\ref{fig:nev2}, in the boosted category
events, with a cut in the ANN output as hard as $y_{\rm cut}\simeq 0.9$,
we are still left
with around 100 events at the HL-LHC, 
which the background being reduced substantially to around
1000 events.
%
A similar cut would work fine for the intermediate category, though here
the backgrounds are much larger and thus the expected signal significance smaller.
%
For the resolved case, the signal significance would be only slightly
improved for any value of the cut.

A useful property of MVAs such as the one used in this work
is that they provide direct  physical insight about which of the
input variables contribute the most to the separation of
signal over background.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of al the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm tot}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
The input variables with a larger value of $\omega^{\rm tot}_i$ will be those
that play a dominant role in enhancing the signal
discrimination using the MVA.

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
The notation for the various kinematic variables is the same
as in Sect.~\ref{sec:input}.
%
The important information
is contained in the relative strengths of the total associated weight
for each of the input variables.
%
The results of Fig.~\ref{fig:nnweights} illustrate which variables
are more discriminative between the signal and the background.
%
For instance, in the 
resolved category, the variables that lead to
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs candidates, and then on a similar
footing other variables like the $p_T$ of the individual jets
and the mass of the di-Higgs system $m_{hh}$.
%
The only variable which seems to contain no information at all
is the $\Delta \phi_{hh}$ separation between the two
Higgs candidates.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/res_wgthist_noPU.pdf}
\includegraphics[width=0.49\textwidth]{plots/int_wgthist_noPU.pdf}
\includegraphics[width=0.75\textwidth]{plots/bst_wgthist_noPU.pdf}
\vspace{-0.5cm}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (upper left plot),  intermediate (upper right plot)
and boosted (lower plot)
categories.
%
See Sect.~\ref{sec:input} for a description of each of the input
variables to the MVA.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

For the boosted category, we see that substructure variables
are the most helpful ones, in particular the $k_t$-splitting scale
$\sqrt{d_{12}}$ and the subjettiness ratio $\tau_{12}$.
%
The ratios of ECFs $D_2^{(\beta)}$ and
$C_2^{(\beta)}$ also seems to provide
useful discrimination power.
%
Concerning other kinematic variables, the angular
separation between the two large-$R$ jets $\Delta R_{hh}$,
the invariant mass of the di-Higgs system $m_{hh}$ and the various
$p_T$ also have discrimination power.
%
Other variables provide limited information, in particular the invariant
mass of the Higgs candidates, $m_{h_1}$ and $m_{h_2}$: this
is due to the realistic
simulation of detector resolution introduced with the smearing.
%
We have verified that the post-MVA
signal significance in the boosted category is substantially
degraded if substructure variables are not used.

At this point we have all the
information needed to determine the optimal
value of the MVA output cuts $y_{\rm cut}$
and quantify the final values of the signal
significance $S/\sqrt{B}$ and $S/B$ in the three categories.
%
These optimal values are determined from the maximisation of $S/\sqrt{B}$,
ensuring that the number of events $N_{\rm ev}$
which would be expected at the HL-LHC is large
enough.
%
In addition, we also require when fixing $y_{\rm cut}$
that the actual number of MC events used for the ANN training
is large enough to avoid the biases of a small training sample.

These results are summarized in Table~\ref{table:cutflowMVA}, where we indicate
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC,
and also $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 1070 & $7.6\cdot 10^5$  & 1.2  & $1.4\cdot 10^{-3}$  \\
    &  $y_{\rm cut}=0.82$ & 790  & $2.2\cdot 10^4$   & 5.4  & 0.034 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  & 670   & $5.3\cdot 10^5$
    & 0.9 & $1.2\cdot 10^{-3}$ \\
    &  $y_{\rm cut}=0.80$ & 360  & $5.5\cdot 10^3$  & 4.8 & 0.06\\
    \hline
    \hline
      \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  & 3700 &  $1.5\cdot 10^{7}$ &  1.0 &$3\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.65$ & 1300  & $8.3\cdot 10^{4}$ & 4.5 & 0.01 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA results, for the optimal value of the
    ANN discriminant $y_{\rm cut}$ in the three categories, compared with the
    corresponding
    pre-MVA results $y_{\rm cut}=0$.
    %
    We indicate the number of signal and
    background events
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    the signal significance $S/\sqrt{B}$ and
    the signal over background ratio $S/B$.
    %
    The pre-MVA results ($y_{\rm cut}=0$) corresponds to row C2 in
    Table~\ref{tab:cutflow_noPU_1}.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The post-MVA results for $S/\sqrt{B}$ and $S/B$ as a function of the cut
in the ANN output for each of the three categories are shown in
Fig.~\ref{fig:sb_mva}.
%
Note that the values of $S/\sqrt{B}$ and $S/B$
for $y_{\rm cut}=0$ correspond to the values at
the end of the cut-based analysis, see
the C2 row of Table~\ref{tab:cutflow_noPU_1}.
%
Indeed, at the end of the cut based analysis, the
signal significance $S/\sqrt{B}$ was
1.2, 0.9 and 1.0 for the boosted, intermediate and resolved
categories respectively, consistent with Fig.~\ref{fig:sb_mva}.
%
The same applies to the signal over background ratio $S/B$.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_noPU.pdf}
\caption{\small
  The values of the signal significance $S/\sqrt{B}$ and of the
  signal over background ratio $S/B$ for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}=0$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{tab:cutflow_noPU_1}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

The results of Fig.~\ref{fig:sb_mva}  show that
for all categories the  improvement in signal
significance as compared to the pre-MVA results
is substantial, growing from around 1
to around 5 in the three cases.
%
The optimal values of $y_{\rm cut}$ ensure that the
number of events at the HL-LHC is still well appreciable: we obtain
1300, 360 and 800 events in the resolved, intermediate, and
boosted categories respectively.
%
Another important benefit of the MVA applied to the boosted
category is that fact that the background is substantially
reduced so that $S/B$ can be as large as 10\% for the optimal
values of the ANN output cut.
%
Note that this is not true for the other two categories,
where $S/B$ is roughly constant.
%
A reasonably large value of $S/B$ is beneficial
from the experimental point of view, since it indicates
that the measurement is feasible even if the
systematic uncertainties are not very small.

From Table~\ref{table:cutflowMVA} we see that after the
MVA the signal significance in the boosted category increases
from 0.99 to 3.31, with $S/B$ increasing from $0.005$ to $0.10$,
with still more that 100 signal events at the HL-LHC.
%
Table~\ref{table:cutflowMVA} also shows that for the intermediate
and resolved categories, the improvement due to the use
of the MVA is rather moderate, and signal significance is always
small since backgrounds are still  large.
%
The combination of the signal significance of the
three categories leads to a $S/\sqrt{B}=3.53$, and compared
to 3.31 using only the boosted category.

Table~\ref{fig:sb_mva} contains the main result of this paper:
at the HL-LHC, it should be possible to claim evidence of
Higgs pair production using only the $4b$ final state.
%
Needless to say, to claim discovery and to improve the accuracy
of the measurement of the trilinear coupling it would still
be needed to combine this channel with others like
$b\bar{b}\gamma\gamma$ and $b\bar{b}\tau\tau$, but in any case
it is always important to have robust evidence from
individual channels.

These remarkable results are however subject to an obvious criticism:
at the HL-LHC the intense PU is know to substantially affect
kinematical distributions, and in particular, it is expected to degrade
in some extent the results of the present analysis.
%
To preempt this criticism, in the next section
we discuss how the expectations for the HL-LHC
summarized in  Table~\ref{fig:sb_mva} are modified in the presence
of PU expected at the HL-LHC.
%
But before that let us illustrate how the improvement in $S/\sqrt{B}$
obtained from the MVA in the boosted
category arises from the information contained in the
jet substructure variables.

\subsection{Another look at jet substructure variables}

%
From the distribution of total associated ANN weights in the
boosted category, shown in Fig.~\ref{fig:nnweights}, we see that
all substructure variables contain useful information:
the $k_t$-splitting scales $\sqrt{d_{12}}$,
the $\tau_{21}$ subjettiness ratio, and the ECF ratios
$C_2^{(\beta)}$ and $D_2^{(\beta)}$.
%
Interestingly, for some substructure variables, such as $\tau_{21}$,
the total associated weight turns out to be stronger in the leading
than in the sub-leading Higgs candidate (or vice-versa), indicating some
degree of redundancy when the same variable is used for the two Higgs candidates
at the same time.

The information reflected in Fig.~\ref{fig:nnweights} should also be
apparent from the corresponding distributions of these substructure
variables: those with a higher discrimination power should show
clear differences in shape between the signal and background events,
while those that carry less information should have similar
shapes for signal and background.
%
The third possibility is that of of redundant variables, where signal
and background exhibit different shapes but the same information
is already provided by other input variables.
%
To illustrate these points,
in Fig.~\ref{fig:mva_substructure_1}
we show the distributions of some of
these substructure variables for the boosted category: the
$k_t$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}), for
the subleading Higgs,
the energy correlation double ratio $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}), for the leading Higgs,
and then 
the subjettiness ratio $\tau_{12}$,
Eq.~(\ref{eq:tau21}),
for both the leading and subleading Higgs candidates.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{plots/split12_h1_C1_boost.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h1_C1_boost.pdf}
  \caption{\small Distribution of substructure variables
    in the boosted category at the end of the cut-based
    analysis, used as input to the MVA.
    %
    From top to bottom and from left to right we show  the
$k_t$ splitting scale for
the subleading Higgs, the subjettiness ratio $\tau_{12}$
for the leading Higgs,
and the energy correlation double ratio $D_2$
for the leading and subleading Higgs.
%
All distributions have been normalized to unity.
}
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Fig.~\ref{fig:mva_substructure_1}
we can observe that for these substructure variables the shapes of the signal
and background distributions show important differences,
reflecting the inherent differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.
%
Importantly, we observe that signal and background distributions peak
in different regions: for example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (50 GeV) for signal (background) events, while
the ECF ratio $D_2^{(\beta)}$ peaks at much lower values for signal than
for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are quite similar
in the leading and in the subleading jets, therefore adding
$\tau_{12}$ from the two Higgs candidates to the MVA at the same time
is redundant, explaining why in these cases either one or the other
variable from a Higgs candidate is assigned higher weight.

It is also interesting to show the distributions of kinematic variables
that according to the MVA provide a small amount of discrimination
between signal and background, and where the should not be an issue
of redundancy.
%
In Fig.~\ref{fig:mva_substructure_2}
we show the kinematic variable that carries less weight in the MVA,
    the azimuthal angle separation between the two Higgs candidates
    $\Delta\phi_{hh}$, for the resolved and boosted categories:
    according to Fig.~\ref{fig:nnweights}, essentially no discrimination power
    should be obtained from this variable, and indeed the shapes of the two signal
    and background
    are very similar here.
    %
    From Fig.~\ref{fig:nnweights} we also saw that the
    invariant mass distributions $m_h$ carry little information
    to discriminate signal over background:
    this was already verified from 
the corresponding
distributions in Fig.~\ref{fig:mHHinv}.
%
As explained before, the reduced discrimination power arises
from 
the  smearing of jet four-momenta introduced to mimic
realistically detector resolution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.48\textwidth]{plots/DeltaPhi_HH_C1_res.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/DeltaPhi_HH_C1_boosted.pdf} 
  \caption{\small Same as Fig.~\ref{fig:mva_substructure_1},
    now for the kinematic variable that carries less weight in the MVA,
    the azimuthal angle separation between the two Higgs candidates
    $\Delta\phi_{hh}$, for the resolved and boosted categories.
}
\label{fig:mva_substructure_2}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Impact of PU}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 360   &  $1.4\cdot 10^6$ & 0.31   &
     $3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.82$ &  250 & 7300  & 2.9    & 0.034  \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  &  150  & $9\cdot 10^5$    & 0.15    &
     $2\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.80$ & 50 & 2600  &  1.1   & 0.02 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  &  1100  & $8.5\cdot 10^6$
    & 0.25    &  $1.3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.65$ & 430  & $3\cdot 10^4$  &  2.5   & 0.15  \\
    \hline
      \end{tabular}
  \caption{\small Same as Table~\ref{table:cutflowMVA}, now for the case
    of PU+SK, with $\la n_{\rm PU}\ra=80$.
        \label{table:cutflowMVA_PU}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

It important to verify, for each value of
the cut in the ANN output $y_{\rm cut}$, how many
signal and background events are expected at the HL-LHC,
with an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
%
This comparison is shown in 
Fig.~\ref{fig:nev2}.
%
As can be seen,
in the boosted category, for a value $y_{\rm cut}\simeq 0.8$
we end up with 800 signal events and around $2\cdot 10^4$ background
events.
%
Similar results are obtained in the intermediate and resolved
categories: in the former we find 700 ($6\cdot 10^3$) signal (background)
events for $y_{\rm cut}\simeq 0.8$, and in the latter
1300 ($8\cdot 10^4$) signal (background) events for
$y_{\rm cut}\simeq 0.65$.
%
Note that the MVA achieves a very substantial background suppression
with only a
moderate reduction of signal efficiency in exchange.

A useful property of MVAs such as the one used in this work
is that they provide direct  physical insight about which of the
input variables contribute the most to the separation between
signal and background.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of all the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm (tot)}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
Those input variables with a larger value of $\omega^{\rm (tot)}_i$ will be those
that play a dominant role in enhancing the signal
discrimination using the MVA.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/res_wgthist_noPU.pdf}
\includegraphics[width=0.49\textwidth]{plots/int_wgthist_noPU.pdf}
\includegraphics[width=0.75\textwidth]{plots/bst_wgthist_noPU.pdf}
\vspace{-0.5cm}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (upper left),  intermediate (upper right)
and boosted (lower plot)
categories.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the three categories, using the
notation for the various kinematic variables
as in Sect.~\ref{sec:input}.
%
The important information
is contained in the relative strengths of the total associated weight
for each of the input variables.
%
In the 
resolved category, the variables that carry 
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs candidates and
their invariant masses $m_{h1}$ and $m_{h2}$.
%
In the case of the boosted category, the invariant mass distribution
of the Higgs candidates is also the most discriminatory
variable, followed by the subjet $p_T$ distributions and
substructure variables such as $C_2^{(\beta)}$ and
$D_2^{(\beta)}$.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_noPU.pdf}
\caption{\small
  The values of the signal significance, $S/\sqrt{B}$, and of the
  signal over background ratio, $S/B$, for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{tab:cutflow_noPU_1}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

The results for the signal significance $S/\sqrt{B}$ and
the signal over background ratio
$S/B$ as a function of $y_{\rm cut}$
for the three categories are
Fig.~\ref{fig:sb_mva}.
%
The values 
for $y_{\rm cut}=0$ correspond to those
the end of the cut-based analysis, see
Table~\ref{tab:cutflow_noPU_1}.
%
We observe how in the three
 categories there is a marked  improvement in signal
significance as compared to the pre-MVA results.
%
We also observe a substantial enhancement in $S/B$, arising
from the background suppression indeed by the MVA, reaching
values of 1\%, 6\% and 3.5\% in the resolved,
intermediate and boosted categories.
%
This improvement in $S/B$ is crucial to ensure the feasibility
of this measurement, since it allows systematic
uncertainties in the background determination to
be at the same level of a few percent level.

We now have all the
information needed to determine a suitable
value of the MVA output cut $y_{\rm cut}$ in each
of the three categories.
%
These values can be determined from the maximisation of $S/\sqrt{B}$,
ensuring that the number of signal events $N_{\rm ev}$
expected at the HL-LHC is not too small.
%
In  addition, we require
that the number of MC events used to define the signal
category (events with $y_i \ge y_{\rm cut}$)
is large enough to avoid the biases and statistical
fluctuations associated to a small training sample.
%
In Table~\ref{table:cutflowMVA} where we indicate
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC as well as $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
    \hline
    \multicolumn{6}{|c|}{No PU} \\
    \hline
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 1070 & $7.6\cdot 10^5$  & 1.2  & $1.4\cdot 10^{-3}$  \\
    &  $y_{\rm cut}=0.82$ & 790  & $2.2\cdot 10^4$   & 5.4  & 0.034 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  & 670   & $5.3\cdot 10^5$
    & 0.9 & $1.2\cdot 10^{-3}$ \\
    &  $y_{\rm cut}=0.80$ & 360  & $5.5\cdot 10^3$  & 4.8 & 0.06\\
    \hline
    \hline
      \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  & 3700 &  $1.5\cdot 10^{7}$ &  1.0 &$3\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.65$ & 1300  & $8.3\cdot 10^{4}$ & 4.5 & 0.01 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA results, for the optimal value of the
    ANN discriminant $y_{\rm cut}$ in the three categories, compared with the
    corresponding
    pre-MVA results $y_{\rm cut}=0$.
    %
    We indicate the number of signal and
    background events
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    the signal significance $S/\sqrt{B}$ and
    the signal over background ratio $S/B$.
    %
    The pre-MVA results ($y_{\rm cut}=0$) corresponds to row C2 in
    Table~\ref{tab:cutflow_noPU_1}.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Table~\ref{table:cutflowMVA} we see that at the end of the
MVA the signal significance in the boosted category increases
from 1.2 to 5.4, with $S/B$ increasing from $0.14\%$ to $3.4\%$,
with almost 800 signal events expected at the HL-LHC.
%
For the intermediate and resolved category, $S/\sqrt{B}$
increases from 0.9 and 1.0 to 4.8 and 4.5, with
the signal over background ratio raising from
$0.12\%$ and $0.03\%$ to 6\% and 1\$, respectively.
%
The overall combined significance is thus $S\sqrt{B}\simeq 8$,
well above the threshold for discovery.
%
These remarkable results are however subject to a justified
criticism:
the HL-LHC will be a high-PU environment,
which will affect the description of the various
kinematical distributions used as input to the MVA.
%
Therefore, next we quantify the robustness of the
results summarized Table~\ref{fig:sb_mva}
in a realistic high-PU environment.

\subsection{Impact of PU in the MVA}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
     \multicolumn{6}{|c|}{$\la n_{\rm PU}\ra=80$+SK} \\
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 360   &  $1.4\cdot 10^6$ & 0.31   &
     $3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.82$ &  250 & 7300  & 2.9    & 0.034  \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  &  150  & $9\cdot 10^5$    & 0.15    &
     $2\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.80$ & 50 & 2600  &  1.1   & 0.02 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  &  1100  & $8.5\cdot 10^6$
    & 0.25    &  $1.3\cdot 10^{-4}$  \\
    &  $y_{\rm cut}=0.65$ & 430  & $3\cdot 10^4$  &  2.5   & 0.15  \\
    \hline
      \end{tabular}
  \caption{\small Same as Table~\ref{table:cutflowMVA}, now for the case
    of PU+SK, with $\la n_{\rm PU}\ra=80$.
        \label{table:cutflowMVA_PU}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
